---
title: "Volcano plot"
author: "Anne-Maud Ferreira"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Volcano plot}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  ---
  
  ```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  ```
  
  ```{r setup, warning = FALSE, message = FALSE}
  library(CytoGLMM)
library(SummarizedExperiment)
library(magrittr)
library(tidyverse)
```

# Introduction

This is a short vignette showing an example of the volcano plot.

# Aghaeepour et al. (2017) dataset

As an example, the Aghaeepour et al. (2017) dataset was used:

- Selection of 16 women from the Aghaeepour et al. (2017) dataset 
- Samples from the first and third trimester of their pregnancy stimulated by IFN$\alpha$


```{r load_data}
# Path
zenodo_url <- "https://zenodo.org/record/2652578/files/"
cytof_data <- "se_aghaeepour2017immune.Rdata"
# Download data
## se_aghaeepour2017immune is an SummarizedExperiment object
load(url(paste0(zenodo_url, cytof_data)))
```

## Pre-processing of the data 

The data are processed as in the example of the CytoGLMM workflow available online: 

```{r process_data}
# Extract the values
exprs <- assay(se_aghaeepour2017immune)
# Extract the sample information
sample_info <- rowData(se_aghaeepour2017immune)
# Extract the sample names
sample_info_names <- names(sample_info)
# Create a data frame with the expression values and the sample information
df_samples <- cbind(as.data.frame(exprs), as.data.frame(sample_info))
df_samples %<>% as_tibble
# Extract protein names
protein_names <- colData(se_aghaeepour2017immune) %>% 
  as.data.frame %>% 
  dplyr::filter(type == "function") %>%
  .$protein_name

# Subset to NK cell population
df_samples_subset <- df_samples %>% dplyr::filter(celltype == "NK")
df_samples_subset %<>% dplyr::select(protein_names, sample_info_names)

# Transform the data
df_samples_subset_transformed <- df_samples_subset %>% 
  dplyr::mutate_at(protein_names, 
            function(x) asinh(x/5))


```

## Fit the GLMM model

Fit a Generalized Linear Mixed Model (GLMM) with donor random effects.

```{r fit_cytoglmm}
glmm_fit <- CytoGLMM::cytoglmm(df_samples_subset_transformed, 
                               protein_names = protein_names,
                               condition = "term", group = "donor")
```

# Volcano plot

Plot the volcano plot as a complementary plot for the CytoGLMM model output.

```{r volcano_plot}
# Prepare the data
data_for_volplot <- prepare_data_for_volcanoplot(data = df_samples_subset, 
                                                 protein_name = protein_names,
                                                 condition = "term", 
                                                 CytoGLMM_fit = glmm_fit)
# Volcano plot
volcano_plot(data = data_for_volplot, 
             exp_conditions = c("1st trim", "3rd trim"))
```

